---
title: "Why Move? Finding State Attractiveness"
subtitle: "Project 2"
author: "James Aas, Audrey Mills, Jameson Oates"
format: html
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r load-packages}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(dplyr)
library(tidyr)
library(tmap)
library(readr)
library(leaflet)
library(maps)
library(gganimate)
library(moderndive)
library(ggplot2)
library(shiny)
```

```{r load-data}
#| label: load-data
#| message: false
#| warning: false

state_economic_data <- read.csv("data/state_economic_data.csv")

state_abortion_data <- read.csv("data/abortions_dataset.csv")

state_immigration_data <- read.csv("data/interstate_migration_data.csv")

presidential_election_data <- read.csv("data/presidential_election_data.csv")

state <- c("Alabama", 
           "Alaska", 
           "Arizona", 
           "Arkansas",
           "California",
           "Colorado",
           "Connecticut",
           "Delaware",
           "District of Columbia",
           "Florida",
           "Georgia",
           "Hawaii",
           "Idaho",
           "Illinois",
           "Indiana",
           "Iowa",
           "Kansas",
           "Kentucky",
           "Louisiana",
           "Maine",
           "Maryland",
           "Massachusetts",
           "Michigan", 
           "Minnesota",
           "Mississippi",
           "Missouri",
           "Montana",
           "Nebraska",
           "Nevada",
           "New Hampshire",
           "New Jersey",
           "New Mexico",
           "New York",
           "North Carolina",
           "North Dakota",
           "Ohio",
           "Oklahoma",
           "Oregon",
           "Pennsylvania",
           "Rhode Island",
           "South Carolina",
           "South Dakota",
           "Tennessee",
           "Texas",
           "Utah",
           "Vermont",
           "Virginia",
           "Washington",
           "West Virginia",
           "Wisconsin", 
           "Wyoming")

political_culture <- c("Traditionalistic", 
                       "Individualistic", 
                       "Traditionalistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Individualistic",
                       "Individualistic", 
                       "Individualistic", 
                       "Traditionalistic",
                       "Traditionalistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Individualistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Individualistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic",
                       "Individualistic",
                       "Moralistic",
                       "Individualistic",
                       "Individualistic",
                       "Moralistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic",
                       "Individualistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic", 
                       "Traditionalistic",
                       "Traditionalistic",
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Traditionalistic",
                       "Moralistic", 
                       "Individualistic")

political_culture_elazar <- data.frame(state, political_culture) %>%
  mutate(state = tolower(state))

state_abortion_data <- state_abortion_data %>%
  slice(-(1:1)) %>%
  rename(
	Area = 1,
	Number = 2,
	Rate = 3,
	Ratio = 4,
	`Out-of-State (number and percentage)` = 5,
	Year = 6
  ) %>%
  mutate(
	Number = ifelse(Number %in% c("N/A"), NA, Number),
	Rate = ifelse(Rate %in% c("N/A"), NA, Rate),
	Ratio = ifelse(Ratio %in% c("N/A"), NA, Ratio),
	`Out-of-State (number and percentage)` = ifelse(`Out-of-State (number and percentage)` %in% c("N/A"), NA, `Out-of-State (number and percentage)`)
  )  %>%
  filter(!(Number %in% c("25,619",
                     	"30,555",
                     	"35,398",
                     	"37,523",
                     	"73,815",
                     	"89,469",
                     	"59,854",
                     	"27,471"))) %>%
  mutate(Number = as.numeric(gsub(",", "", Number)),
     	Rate = as.numeric(gsub(",", "", Rate)),
     	Ratio = as.numeric(gsub(",", "", Ratio)),
     	Year = as.numeric(gsub(",", "", Year))
     	) %>%
  mutate(Area = tolower(Area)) %>%
  mutate(Abortions_per_1000_women = Rate) %>%
  mutate(Abortions_per_1000_live_births = Ratio) %>%
  mutate(`Out_of_state_abortions(number and percentage)` = `Out-of-State (number and percentage)`) %>%
  select(Area, Abortions_per_1000_women, Abortions_per_1000_live_births, Year, `Out_of_state_abortions(number and percentage)`)
  

state_immigration_data <- state_immigration_data %>%
   slice(-(1:6)) %>%
  rename(
    `Current residence in`  = 1,
    `Population 1 year and over: Estimate` = 2,
    `Population 1 year and over: MOE` = 3,
    `Same house 1 year ago: Estimate` = 4,
    `Same house 1 year ago: MOE` = 5,
    `Same state of residence 1 year ago: Estimate` = 6,
    `Same state of residence 1 year ago: MOE` = 7,
    `Different state of residence 1 year ago: Estimate` = 8,
    `Different state of residence 1 year ago: MOE` = 9,
    Year = 10
  ) %>%
  slice(-(1:6)) %>%
  mutate(
    `Population 1 year and over: Estimate` = as.numeric(
      gsub(",", "", `Population 1 year and over: Estimate`)
      ),
    `Same house 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Same house 1 year ago: Estimate`)
      ), 
    `Same state of residence 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Same state of residence 1 year ago: Estimate`)
      ), 
    `Different state of residence 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Different state of residence 1 year ago: Estimate`)
      ),
     `Different state of residence 1 year ago: MOE` = as.numeric(
      gsub(",", "", "+/-", `Different state of residence 1 year ago: MOE`)
      ),
     `Same state of residence 1 year ago: MOE` = as.numeric(
      gsub(",", "", "+/-", `Same state of residence 1 year ago: MOE`)
      ),
    `Same house 1 year ago: MOE` = as.numeric(
      gsub(",", "", "+/-", `Same house 1 year ago: MOE`)
      ), 
    `Population 1 year and over: MOE` = as.numeric(
      gsub(",", "", "+/-", `Population 1 year and over: MOE`)
      ), 
    ) 


state_immigration_data<-state_immigration_data %>%
  mutate(Different_state_rate = `Population 1 year and over: Estimate` / `Different state of residence 1 year ago: Estimate`)

state_economic_data <- state_economic_data %>%
  rename(
    State  = 1,
    `Median Household Income` = 2,
    Year = 3
  ) %>%
  mutate(
    State = ifelse(State == "", NA, State),
    `Median Household Income` = ifelse(`Median Household Income` == "", NA, `Median Household Income`)
    ) %>%
  filter(complete.cases(.)) %>%
  mutate(`Median Household Income` = as.numeric(gsub(",", "", `Median Household Income`))) %>%
  mutate(State = tolower(State))

presidential_election_data <- presidential_election_data %>%
  subset(select = -c(X1, X, TOTAL.VOTES, TOTAL.VOTES..)) %>%
  mutate(
    FEC.ID = ifelse(FEC.ID == "n/a", NA, FEC.ID),
    `GENERAL.RESULTS` = as.numeric(gsub(",", "", `GENERAL.RESULTS`)),
    STATE = tolower(STATE),
    `R Share of General` = as.numeric(gsub("%", "", `GENERAL..`)) / 100
  ) %>%
  filter(PARTY == "R") %>%
  select(STATE, `R Share of General`, YEAR)

abortion_and_economic <- full_join(state_abortion_data, 
             state_economic_data,
             by = c("Area" = "State",
                    "Year" = "Year"))
abortion_and_economic <- abortion_and_economic %>%
  mutate(Year = as.numeric(gsub(",", "", Year)))

state_immigration_data <- state_immigration_data %>%
  mutate(Year = as.numeric(gsub(",", "", Year))) 

culture_president <- left_join(presidential_election_data, 
                               political_culture_elazar,
                                 by = c("STATE"= "state")) 
everything_but_immigration <- left_join(culture_president, 
                               abortion_and_economic,
                    by = c("STATE" = "Area",
                    "YEAR" = "Year"))

all_variables_dataset<- full_join(everything_but_immigration, 
            state_immigration_data,
            by = c("STATE" = "Current residence in",
                   "YEAR" = "Year"))
```

## **ABSTRACT**

This project gathers data from online sources regarding various cultural, political, and social factors and uses it to create multiple spacial-temporal visualization illustrating the variation in the selected factors between states, with the goal of identifying patters in state-to-state migration.

## **INTRODUCTION**

The variables examined here are state abortion policy, economic status, political culture, and presidential vote in select years.

The dataset "state_immigration_data," from the US Census, describes the population estimates and margin of errors of populations, per state and year, of people living in the state one year ago, living in the state in the same house one year ago, living in a different state one year ago, and total population. Recorded are the years 2012, 2016, and 2020. This dataset lets us track the population levels of each state over time, based on who lived where one year before being recorded.

The dataset "state_abortion_data," from the CDC, contains information on the number of abortions per 1000 women in each state in the years 2008, 2012, 2016, and 2020. It also includes information on abortions per 1000 live births and the number of out-of-state abortions. This lets us visualize an important social factor that might motivate people to migrate in or out of their state of residence.

The dataset "presidential_election_data," from the Federal Election Commission, contains the share of Republican votes in each state per election in 2008, 2012, and 2020. This lets us visualize the political demographics in states, which might motivate interstate migration.

The dataset "state_economic_data," from the Federal Reserve, contains information on median household income for each state in the years 2008, 2012, 2016, and 2020. This lets us see if there might be economic reasons for interstate migration, in correlation with the other datasets.

The dataset "political_culture_elazar" correlates each state with a culture tag, as theorized by Daniel Elazar in 1966. This will allow us to recognize patterns between state culture and the other variables of our analysis.

The other datasets combine the above variables of interest in different ways for the purpose of building a predictive model.

## VISUALIZATIONS

```{r}
econ_data <- state_economic_data

econ_map_data <- map_data("state")

econ_merged_data <- merge(econ_map_data, econ_data, by.x = "region", by.y = "State", all.x = TRUE)

econ_animate <- ggplot(econ_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = `Median Household Income`)) +
  scale_fill_gradient(name = "Income in USD", low = "white", high = "darkblue", na.value = "grey", guide = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "cm")) +
  labs(subtitle = "Real Median Household Income over Time", caption = "Source: FRED") +
  transition_states(Year) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(econ_animate, renderer = gifski_renderer())
```

```{r}
abortion_data <- state_abortion_data

abortion_map_data <- map_data("state")

abortion_merged_data <- merge(abortion_map_data, abortion_data, by.x = "region", by.y = "Area", all.x = TRUE)

abortion_animate <- ggplot(abortion_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Abortions_per_1000_women)) +
  scale_fill_gradient(name = "Abortions per 1000 Women", low = "white", high = "darkgreen", na.value = "grey", guide = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "cm")) +
  labs(subtitle = "Abortions per 1000 Women Aged 15-44 Over Time", caption = "Source: CDC") +
  transition_states(Year) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))

animate(abortion_animate, renderer = gifski_renderer())
```

```{r}
election_data <- presidential_election_data

election_map_data <- map_data("state")

election_merged_data <- merge(election_map_data, election_data, by.x = "region", by.y = "STATE", all.x = TRUE)

election_animate <- ggplot(election_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = `R Share of General`)) +
  scale_fill_gradient(name = "Percent of Vote Republican", low = "white", high = "darkred", na.value = "grey", guide = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "cm")) +
  labs(subtitle = "Republican Share of Vote By State in General Presidential Election", caption = "Source: FEC") +
  transition_states(YEAR) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(election_animate, renderer = gifski_renderer())
```

```{r}
election_merged_data <- merge(election_map_data, election_data, by.x = "region", by.y = "STATE", all.x = TRUE)

election_animate <- ggplot(election_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = (100 - `R Share of General`) - 99)) +
  scale_fill_gradient(name = "Percent of Vote Democrat", low = "white", high = "darkblue", na.value = "grey", guide = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "cm")) +
  labs(subtitle = "Democrat Share of Vote By State in General Presidential Election", caption = "Source: FEC") +
  transition_states(YEAR) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(election_animate, renderer = gifski_renderer())
```

```{r}
political_data <- political_culture_elazar

political_map_data <- map_data("state")

political_merged_data <- merge(political_map_data, political_data, by.x = "region", by.y = "state", all.x = TRUE)

political_map <- ggplot(political_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = political_culture), color = "black", size = 0.2) +
  scale_fill_manual(values = c("lightgreen", "magenta", "lightyellow"), name = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "cm")) + 
  labs(subtitle = "Elazar's Cultural Classification by State", caption = "Source: Daniel Elazar")  

political_map
```

```{r}
state_migration_data <- state_immigration_data

state_migration_map_data <- map_data("state")

state_migration_merged_data <- merge(state_migration_map_data, state_migration_data, by.x = "region", by.y = "Current residence in", na.rm= TRUE)

migration_animate <- ggplot(state_migration_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = `Different_state_rate`)) +
  scale_fill_gradient(name = "Percent of New Residents", low = "white", high = "darkgreen", na.value = "grey", guide = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "cm")) +
  labs(subtitle = "Estimated percentage of state population over age 1 who lived in a different state the previous year over time", caption = "Source: Census") +
  transition_states(Year) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(migration_animate, renderer = gifski_renderer())
```

## **LINEAR MODEL**

```{r}
editing <- all_variables_dataset %>% 
  filter(YEAR == '2008')

all_variables_dataset <- anti_join(all_variables_dataset, editing, by = 'YEAR')
```

```{r}
immigration_log_mlr <- lm(log(`Different_state_rate`) ~
                        log(`Median Household Income`) +
                        log(`Abortions_per_1000_women`) + 
                        log(`R Share of General`), 
                      data = all_variables_dataset, na.rm = TRUE)

get_regression_table(immigration_log_mlr)
```

The intercept is 7.398, this means that in a state with a log Median household income of \$0, log 0 abortions per 1000 women, and a 0 percent Republican share of the general election, the expected rate of the population that comes from a different state is 7.398. The log estimates are telling as well, for a increase of 1 in log of median household income, there is an associated increase of -0.367 in the number of people from different states entering the state compared to the population in the state at the time, holding all other variables constant. For log of abortions per 1000 women the estimate is a minimal change, and it has a p_value of 0.554, leading us to not have a lot of faith in the estimate. The Republican share of the general election has a similar impact to the median household income, but is slightly less pronounced. The p_value for log of Median Household Income is closest to 0.05, meaning that our trust in that prediction is the highest.

```{r}
immigration_lm <- lm(`Different_state_rate` ~
                        `political_culture`, 
                      data = all_variables_dataset, na.rm = TRUE
  )

get_regression_table(immigration_lm)

```

This model was constructed due to the other model being unable to account for categorical variables. It is possible that some sort of dummy variable system could be constructed to include the variables in the logarithmic model, but that was not the method we settled on. The model uses a baseline of individualistic culture. From this baseline we can observe the two conditions. The differences from the baseline are telling, the moralistic cultures have the lowest predicted rate of people from different states moving in, while the traditionalistic states are very similar. Individualistic states are far and above the other two, and has a p-value that leads us to trust the result more than for the other two.

## INTERACTIVE DASHBOARD

Link to access the dashboard: <http://127.0.0.1:4441/>

```{r}
all_variables_dataset <- subset(all_variables_dataset, YEAR != 2016)

usMap <- function(dataset, variable) {
  us_map_data <- map_data("state")
  merged_data <- merge(us_map_data, dataset, by.x = "region", by.y = "STATE", all.x = TRUE)

  color_gradients <- list(
    "R Share of General" = c("white", "deeppink4"), 
    "Median Household Income" = c("white", "darkcyan"), 
    "Abortions_per_1000_women" = c("white", "chartreuse4"),
    "Abortions_per_1000_live_births" = c("white", "chartreuse4"),
    "Same house 1 year ago: Estimate" = c("white", "chocolate4"),
    "Population 1 year and over: Estimate" = c("white", "chocolate4"),
    "Same state of residence 1 year ago: Estimate" = c("white", "chocolate4"),
    "Different state of residence 1 year ago: Estimate" = c("white", "chocolate4")
  )
  
  gradient_colors <- color_gradients[[variable]]
  
  g <- ggplot(merged_data) +
    geom_polygon(aes(x = long, y = lat, group = group, fill = .data[[variable]]),
                 color = "gray70", size = 0.5) +
    scale_fill_gradient(
      low = gradient_colors[1], 
      high = gradient_colors[2],
      na.value = "grey",
      limits = range(dataset[[variable]])
    ) +
    theme_void() +
    theme(legend.position = "bottom") +
    theme(legend.key.width = unit(2, "cm"))
  
  return(g)
}

ui <- fluidPage(
  titlePanel("Why Move? Finding State Attractiveness"),
  mainPanel(
    selectInput("variable", "Variable:", 
                choices = colnames(all_variables_dataset)[!colnames(all_variables_dataset) %in% c("STATE", "YEAR", "Out_of_state_abortions(number and percentage)", "Different state of residence 1 year ago: MOE", "Same state of residence 1 year ago: MOE", "Same house 1 year ago: MOE", "Population 1 year and over: MOE", "political_culture")]),
    selectInput("year", "Year:", choices = unique(all_variables_dataset$YEAR)),
    plotOutput("us_map", width = "750px", height = "450px")
  )
)

server <- function(input, output) {
  output$us_map <- renderPlot({
    subset_data <- all_variables_dataset[all_variables_dataset$YEAR == input$year, ]
    
    if (nrow(subset_data) == 0) {
      return("No data available for the selected year.")
    }
    
    usMap(subset_data, input$variable)
  })
}

shinyApp(ui = ui, server = server)
```
