---
title: "The Attractiveness of Cities"
subtitle: "Project 2"
author: "James Aas, Audrey Mills, Jameson Oates"
title-slide-attributes:
  data-slide-number: none
format: revealjs
editor: visual
execute:
  echo: false
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
```

```{r}
#| label: setup
#| include: false

# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r}
#| label: load-data
#| include: false
```

---
title: "Holiday in Film"
subtitle: "A Data Analysis of Holiday Movies"
format: 
  revealjs
highlight-style: a11y
execute:
  code-link: true
  echo: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
#document setup
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo = FALSE}
#load necessary libraries
library(tidyr)
library(readr)
library(ggplot2)
library(dplyr)
library(moderndive)
```

```{r, echo = FALSE}
#acquires data
holiday_movies <- read_csv("data/holiday_movies.csv")

#combines holidays into one column into new data frame, new_holiday_movies
new_holiday_movies <- pivot_longer(
  holiday_movies,
  cols = c("christmas", 
           "hanukkah", 
           "kwanzaa", 
           "holiday"),
  names_to = "holidays",
  values_to = "true/false"
)

#filters only the movies where the associated holiday variable is "TRUE" (eg. if film has "christmas", it will not display "FALSE" for hanukkah, kwanzaa, and holiday). Creates data frame filtered_holiday_movies
filtered_holiday_movies <- new_holiday_movies[new_holiday_movies$`true/false`, ]

#converts the "holidays" column into categories "christmas", "hanukkah", "kwanzaa", "holiday", useful for being graphed later
filtered_holiday_movies$holidays <- factor(filtered_holiday_movies$holidays, levels = c("christmas", "hanukkah", "kwanzaa", "holiday"))
```

## Introduction and Goals {.smaller}

::: notes
-   14 variables, 2265 observations
:::

-   Our dataset: Holiday Movies
    -   From TidyTuesday repository
    -   Dimensions: 2265x14
    -   Created by Jo Hardin, pulling from IMDb
    -   Of note: movies titles containing "Holiday", "Christmas", "Hanukkah", or "Kwanzaa"
-   We looked at trends relating to ratings, age, and run time of different holiday movies
    -   We chose this data set because we are interested in:
        -   How does type of holiday affect a movie's rating, and for each holiday, how does the movie's age affect its rating?
        -   How do movie run times vary over time? Have holiday movies gotten longer over time? Shorter?

##  {.smaller}

::: notes
-   How does type of holiday affect a movie's rating, and for each holiday, how does the movie's age affect its rating?
    -   We combined holidays into one column using pivot longer
    -   Our first graph is a histogram showing the frequency of different average ratings by movie type
    -   Kwanzaa highest ratings (sample size)
    -   Hanukkah variance in ratings (also hard to see, given low number of data points)
    -   Christmas and Holiday most frequent
:::

```{r, echo = FALSE}
#visualizes the frequency of ratings by different holiday movies
ggplot(filtered_holiday_movies, aes(x = average_rating, 
                                    fill = holidays)) +
  geom_histogram(linewidth = .1, color = "black") +
  labs(title = "Holiday movies by rating frequency",
       x = "Average rating",
       y = "Frequency",
       fill = "Holiday in title:") +
  theme_minimal()
```

## Question 1 Graph 2 {.smaller}

::: notes
-   Our second graph is a scatterplot with line of best fit showing the average rating over time for each holiday
    -   Decrease in ratings over time
    -   Fairly consistent
    -   Overall ratings for holiday movies have decreased slightly, fairly consistent
    -   Lack of datapoints for both Kwanzaa and Hanukkah movies
    -   Type of holiday mentioned in the title has a limited effect on the average rating
    -   As year increases the ratings for all types have slightly decreased
:::

```{r, echo = FALSE}
#visualizes the change in ratings over time, faceted by holiday
ggplot(filtered_holiday_movies, aes(x = year, 
                                    y = average_rating, 
                                    color = holidays)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method="lm", 
              se = FALSE, 
              color = "gray36") +
  facet_wrap(~ holidays, labeller = labeller(holidays = c("christmas" = "Christmas", 
                                                          "hanukkah" = "Hanukkah", 
                                                          "kwanzaa" = "Kwanzaa",
                                                          "holiday" = "Holiday"))) +
  labs(title = "Average movie rating by year and holiday",
       x = "Year", 
       y = "Average Rating"
       ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 65, 
                                   vjust = 0.6))
```

## Question 2 Graph 1 {.smaller}

::: notes
-   How do movie run times vary over time? Have holiday movies gotten longer over time? Shorter? Is there a significant difference of run time over time?
    -   Our first graph is a violin plot showing density of runtimes based on holiday
    -   Kwanzaa movies short runtimes
    -   Hanukkah next
    -   Christmas have longest, concentration around 90 minutes
    -   Holiday movies in general -\> 90 minutes but have a wide range
:::

```{r, echo = FALSE}
#visualizes the density of runtimes based on holiday
library(dplyr)
Christmas <- filtered_holiday_movies %>%
  filter(holidays == "christmas") %>%
  filter(year >= 2000 & year <= 2023)

ggplot(filtered_holiday_movies, aes(x = runtime_minutes, 
                                    y = holidays, 
                                    fill = holidays)) +
  geom_violin() +
  labs(title = "Runtime in holiday movies",
    x = "Runtime (in minutes)",
    y = ""
    ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Question 2 Graph 2 {.smaller}

::: notes
-   Our second graph is a a scatterplot showing runtime of movies over time, faceted by holiday type
    -   Runtime has increased slightly over the years in all but "holiday"
    -   Stayed fairly consistent for holiday movies overall
    -   General takeaways
:::

```{r, echo = FALSE}
#visualize the runtime of movies over time, faceted by holiday type
ggplot(filtered_holiday_movies, aes(x = year, 
                                    y = runtime_minutes, 
                                    color = holidays)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method="lm", 
              se = FALSE, 
              color = "gray36") +
  facet_wrap(~ holidays) +
  labs(title = "Movie Runtime by Year and Holiday",
       x = "Year", 
       y = "Runtime in minutes",
       color = "Holiday in movie title") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 65, 
                                   vjust = 0.6))
```

## Linear Model {.smaller}

```{r, echo = FALSE}
#generate a linear model predicting runtime based on year and holiday 
holidays_lm <- lm(data = filtered_holiday_movies, runtime_minutes ~ year + holidays)
get_regression_table(holidays_lm)
```
