---
title: "The Attractiveness of Cities"
subtitle: "Project 2"
author: "Jameson Oates, James Aas, Audrey Mills"
title-slide-attributes:
  data-slide-number: none
format: revealjs
editor: visual
execute:
  echo: false
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(tidyverse)
library(dplyr)
library(tidyr)
library(leaflet)
library(tmap)
library(readr)
library(leaflet)
library(maps)
library(gganimate)
library(moderndive)
library(ggplot2)
```

```{r}
#| label: setup
#| include: false

# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r}
#| label: load-data
#| include: false
#| 
```

```{r, echo = FALSE}
state_economic_data <- read.csv("data/state_economic_data.csv")

state_abortion_data <- read.csv("data/abortions_dataset.csv")

state_immigration_data <- read.csv("data/interstate_migration_data.csv")

state <- c("Alabama", 
           "Alaska", 
           "Arizona", 
           "Arkansas",
           "California",
           "Colorado",
           "Connecticut",
           "Delaware",
           "District of Columbia",
           "Florida",
           "Georgia",
           "Hawaii",
           "Idaho",
           "Illinois",
           "Indiana",
           "Iowa",
           "Kansas",
           "Kentucky",
           "Louisiana",
           "Maine",
           "Maryland",
           "Massachusetts",
           "Michigan", 
           "Minnesota",
           "Mississippi",
           "Missouri",
           "Montana",
           "Nebraska",
           "Nevada",
           "New Hampshire",
           "New Jersey",
           "New Mexico",
           "New York",
           "North Carolina",
           "North Dakota",
           "Ohio",
           "Oklahoma",
           "Oregon",
           "Pennsylvania",
           "Rhode Island",
           "South Carolina",
           "South Dakota",
           "Tennessee",
           "Texas",
           "Utah",
           "Vermont",
           "Virginia",
           "Washington",
           "West Virginia",
           "Wisconsin", 
           "Wyoming")

political_culture <- c("Traditionalistic", 
                       "Individualistic", 
                       "Traditionalistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Individualistic",
                       "Individualistic", 
                       "Individualistic", 
                       "Traditionalistic",
                       "Traditionalistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Individualistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Individualistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic",
                       "Individualistic",
                       "Moralistic",
                       "Individualistic",
                       "Individualistic",
                       "Moralistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic",
                       "Individualistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic", 
                       "Traditionalistic",
                       "Traditionalistic",
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Traditionalistic",
                       "Moralistic", 
                       "Individualistic")

political_culture_elazar <- data.frame(state, political_culture) %>%
  mutate(state = tolower(state))

state_abortion_data <- state_abortion_data %>%
  slice(-(1:1)) %>%
  rename(
	Area = 1,
	Number = 2,
	Rate = 3,
	Ratio = 4,
	`Out-of-State (number and percentage)` = 5,
	Year = 6
  ) %>%
  mutate(
	Number = ifelse(Number %in% c("N/A"), NA, Number),
	Rate = ifelse(Rate %in% c("N/A"), NA, Rate),
	Ratio = ifelse(Ratio %in% c("N/A"), NA, Ratio),
	`Out-of-State (number and percentage)` = ifelse(`Out-of-State (number and percentage)` %in% c("N/A"), NA, `Out-of-State (number and percentage)`)
  )  %>%
  filter(!(Number %in% c("25,619",
                     	"30,555",
                     	"35,398",
                     	"37,523",
                     	"73,815",
                     	"89,469",
                     	"59,854",
                     	"27,471"))) %>%
  mutate(Number = as.numeric(gsub(",", "", Number)),
     	Rate = as.numeric(gsub(",", "", Rate)),
     	Ratio = as.numeric(gsub(",", "", Ratio)),
     	Year = as.numeric(gsub(",", "", Year))
     	) %>%
  mutate(Area = tolower(Area)) %>%
  mutate(Abortions_per_1000_women = Rate) %>%
  mutate(Abortions_per_1000_live_births = Ratio) %>%
  mutate(`Out_of_state_abortions(number and percentage)` = `Out-of-State (number and percentage)`) %>%
  select(Area, Abortions_per_1000_women, Abortions_per_1000_live_births, Year, `Out_of_state_abortions(number and percentage)`)



state_immigration_data <- state_immigration_data %>%
   slice(-(1:6)) %>%
  rename(
    `Current residence in`  = 1,
    `Population 1 year and over: Estimate` = 2,
    `Population 1 year and over: MOE` = 3,
    `Same house 1 year ago: Estimate` = 4,
    `Same house 1 year ago: MOE` = 5,
    `Same state of residence 1 year ago: Estimate` = 6,
    `Same state of residence 1 year ago: MOE` = 7,
    `Different state of residence 1 year ago: Estimate` = 8,
    `Different state of residence 1 year ago: MOE` = 9,
    Year = 10
  ) %>%
  slice(-(1:6)) %>%
  mutate(
    `Population 1 year and over: Estimate` = as.numeric(
      gsub(",", "", `Population 1 year and over: Estimate`)
      ),
    `Same house 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Same house 1 year ago: Estimate`)
      ), 
    `Same state of residence 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Same state of residence 1 year ago: Estimate`)
      ), 
    `Different state of residence 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Different state of residence 1 year ago: Estimate`)
      )
    ) 
state_economic_data <- state_economic_data %>%
  rename(
    State  = 1,
    `Median Household Income` = 2,
    Year = 3
  ) %>%
  mutate(
    State = ifelse(State == "", NA, State),
    `Median Household Income` = ifelse(`Median Household Income` == "", NA, `Median Household Income`)
    ) %>%
  filter(complete.cases(.)) %>%
  mutate(`Median Household Income` = as.numeric(gsub(",", "", `Median Household Income`))) %>%
  mutate(State = tolower(State))
presidential_election_data <- read.csv("data/presidential_election_data.csv")

presidential_election_data <- presidential_election_data %>%
  subset(select = -c(X1, X, TOTAL.VOTES, TOTAL.VOTES..)) %>%
  mutate(
    FEC.ID = ifelse(FEC.ID == "n/a", NA, FEC.ID),
    `GENERAL.RESULTS` = as.numeric(gsub(",", "", `GENERAL.RESULTS`)),
    STATE = tolower(STATE),
    `R Share of General` = as.numeric(gsub("%", "", `GENERAL..`)) / 100
  ) %>%
  filter(PARTY == "R") %>%
  select(STATE, `R Share of General`, YEAR)

abortion_and_economic <- full_join(state_abortion_data, 
             state_economic_data,
             by = c("Area" = "State",
                    "Year" = "Year"))
abortion_and_economic <- abortion_and_economic %>%
  mutate(Year = as.numeric(gsub(",", "", Year)))

state_immigration_data <- state_immigration_data %>%
  mutate(Year = as.numeric(gsub(",", "", Year))) 

culture_president <- left_join(presidential_election_data, 
                               political_culture_elazar,
                                 by = c("STATE"= "state")) 
everything_but_immigration <- left_join(culture_president, 
                               abortion_and_economic,
                    by = c("STATE" = "Area",
                    "YEAR" = "Year"))

all_variables_dataset<- full_join(everything_but_immigration, 
            state_immigration_data,
            by = c("STATE" = "Current residence in",
                   "YEAR" = "Year"))


```

## Question

How is interstate migration affected by state abortion policy, economic status, political culture, and presidential vote?

------------------------------------------------------------------------

## Goals:

-   Examine the numbers of people moving between states and compare them to potential explaining factors such as:

    -   Access to abortion

    -   Median household income

    -   Elazar's cultural classification

    -   Partisanship

-   Create a spacial-temporal visualization of each of these factors from 2012-2020 as well as a linear model examining state migration as a function of them

------------------------------------------------------------------------

## Datasets

::: panel-tabset
## Abortion

```{r}
head(state_abortion_data, 10)
```

## Economic

```{r}
head(state_economic_data, 10)
```

## Culture

```{r}
head(political_culture_elazar, 10)
```

## Partisanship

```{r}
head(presidential_election_data, 10)
```

## Migration

```{r}
head(state_immigration_data, 10)
```
:::

------------------------------------------------------------------------

## Visualizations & Time Series

::: panel-tabset
## Abortions

```{r}
abortion_data <- state_abortion_data

abortion_map_data <- map_data("state")

abortion_merged_data <- merge(abortion_map_data, abortion_data, by.x = "region", by.y = "Area", all.x = TRUE)

abortion_animate <- ggplot(abortion_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = `Abortions_per_1000_women`)) +
  scale_fill_gradient(name = "Abortions per 1000 Women", low = "lightgreen", high = "darkgreen", na.value = "grey", guide = "legend") +
  theme_void() +
  labs(subtitle = "Abortions per 1000 Women Aged 15-44 Over Time", caption = "Source: CDC") +
  transition_states(Year) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(abortion_animate, renderer = gifski_renderer())
```

## Economics

```{r}
econ_data <- state_economic_data

econ_map_data <- map_data("state")

econ_merged_data <- merge(econ_map_data, econ_data, by.x = "region", by.y = "State", all.x = TRUE)

econ_animate <- ggplot(econ_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = `Median Household Income`)) +
  scale_fill_gradient(name = "Income", low = "lightblue", high = "darkblue", na.value = "grey", guide = "legend") +
  theme_void() +
  labs(subtitle = "Real Median Household Income over Time", caption = "Source: FRED") +
  transition_states(Year) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(econ_animate, renderer = gifski_renderer())
```

## Culture

```{r}
political_data <- political_culture_elazar

political_map_data <- map_data("state")

political_merged_data <- merge(political_map_data, political_data, by.x = "region", by.y = "state", all.x = TRUE)

political_map <- ggplot(political_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = political_culture), color = "black", size = 0.2) +
  scale_fill_manual(values = c("lightgreen", "magenta", "lightyellow"), name = "") +
  theme_void() +
  labs(subtitle = "Elazar's Cultural Classification by State", caption = "Source: Daniel Elazar")  

political_map
```

## Partisanship

```{r}
election_data <- presidential_election_data

election_map_data <- map_data("state")

election_merged_data <- merge(election_map_data, election_data, by.x = "region", by.y = "STATE", all.x = TRUE)

election_animate <- ggplot(election_merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = `R Share of General`)) +
  scale_fill_gradient(name = "% of Vote", low = "pink", high = "darkred", na.value = "grey", guide = "legend") +
  theme_void() +
  labs(subtitle = "Republican Share of Vote By State in General Presidential Election", caption = "Source: FEC") +
  transition_states(YEAR) +
  labs(title = "Year: {closest_state}") + 
  theme(plot.title = element_text(hjust = 0.5))  

animate(election_animate, renderer = gifski_renderer())
```
:::

------------------------------------------------------------------------

## Linear Model

```{r}
editing <- all_variables_dataset %>% 
  filter(YEAR == '2008')

all_variables_dataset <- anti_join(all_variables_dataset, editing, by = 'YEAR')
immigration_mlr <- lm(`Different state of residence 1 year ago: Estimate` ~ 
                        `Median Household Income` * 
                        `Abortions_per_1000_women` * 
                        `R Share of General` + 
                        `political_culture`,
                      data = all_variables_dataset, na.rm = TRUE)

get_regression_table(immigration_mlr)


```

------------------------------------------------------------------------

## Current Thoughts & Future Plans

::: panel-tabset
## Now

We have:

-   Spacial-temporal visualizations for:

    -   abortion ratios

    -   median household income

    -   partisanship

-   Spacial visualization for cultural classification

-   Linear model for interstate migration using the above mentioned explanatory variables with interaction between median household income and abortions

## Later

-   Shiny App

    -   Dashboard with editing to select different variables

-   Linear Model

    -   refining

    -   visualizing important findings

-   Overlaying visualizations with migration data

-   Predictive Modeling
:::
