---
title: "The Attractiveness of Cities"
subtitle: "Project 2"
author: "Jameson Oates, James Aas, Audrey Mills"
title-slide-attributes:
  data-slide-number: none
format: revealjs
editor: visual
execute:
  echo: false
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(tidyverse)
library(dplyr)
library(tidyr)
library(leaflet)
library(tmap)
library(readr)
library(leaflet)
library(maps)
library(gganimate)
```

```{r}
#| label: setup
#| include: false

# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r}
#| label: load-data
#| include: false
#| 
```

---
title: "Holiday in Film"
subtitle: "A Data Analysis of Holiday Movies"
format: 
  revealjs
highlight-style: a11y
execute:
  code-link: true
  echo: true
editor_options: 
  chunk_output_type: console
---

Question: How is interstate migration affected by state abortion policy, economic status, political culture, and presidential vote?

## Goals: 

-   Examine the numbers of people moving between states and compare them to potential explaining factors such as:

    -   Access to abortion

    -   Median household income

    -   Elazar's cultural classification

    -   Partisanship

-   Create a spacial-temporal visualization of each of these factors from 2012-2020 as well as a linear model examining state migration as a function of them

## Datasets

## Visualizations

## Linear Model

## Future Plans

abortions per 1000 women
