---
title: "Why Move?"
subtitle: "Proposal"
authors: "James Aas, Audrey Mills, Jameson Oates"
format: html
editor: visual
---

```{r load-packages}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(dplyr)
library(tidyr)
library(leaflet)
library(tmap)
library(readr)
```

## Datasets

```{r load-data}
#| label: load-data
#| message: false

state_economic_data <- read.csv("data/state_economic_data.csv")

state_abortion_data <- read.csv("data/abortions_dataset.csv")

state_immigration_data <- read.csv("data/interstate_migration_data.csv")

presidential_election_data <- read.csv("data/presidential_election_data.csv")

state <- c("Alabama", 
           "Alaska", 
           "Arizona", 
           "Arkansas",
           "California",
           "Colorado",
           "Connecticut",
           "Delaware",
           "District of Columbia",
           "Florida",
           "Georgia",
           "Hawaii",
           "Idaho",
           "Illinois",
           "Indiana",
           "Iowa",
           "Kansas",
           "Kentucky",
           "Louisiana",
           "Maine",
           "Maryland",
           "Massachusetts",
           "Michigan", 
           "Minnesota",
           "Mississippi",
           "Missouri",
           "Montana",
           "Nebraska",
           "Nevada",
           "New Hampshire",
           "New Jersey",
           "New Mexico",
           "New York",
           "North Carolina",
           "North Dakota",
           "Ohio",
           "Oklahoma",
           "Oregon",
           "Pennsylvania",
           "Rhode Island",
           "South Carolina",
           "South Dakota",
           "Tennessee",
           "Texas",
           "Utah",
           "Vermont",
           "Virginia",
           "Washington",
           "West Virginia",
           "Wisconsin", 
           "Wyoming")

political_culture <- c("Traditionalistic", 
                       "Individualistic", 
                       "Traditionalistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Individualistic",
                       "Individualistic", 
                       "Individualistic", 
                       "Traditionalistic",
                       "Traditionalistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Individualistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Individualistic", 
                       "Individualistic", 
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic",
                       "Individualistic",
                       "Moralistic",
                       "Individualistic",
                       "Individualistic",
                       "Moralistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic",
                       "Individualistic",
                       "Individualistic",
                       "Traditionalistic",
                       "Moralistic", 
                       "Traditionalistic",
                       "Traditionalistic",
                       "Moralistic", 
                       "Moralistic", 
                       "Traditionalistic", 
                       "Moralistic", 
                       "Traditionalistic",
                       "Moralistic", 
                       "Individualistic")

political_culture_elazar <- data.frame(state, political_culture) %>%
  mutate(state = tolower(state))

head(political_culture_elazar)
```

```{r}
state_abortion_data <- state_abortion_data %>%
  slice(-(1:1)) %>%
  rename(
	Area = 1,
	Number = 2,
	Rate = 3,
	Ratio = 4,
	`Out-of-State (number and percentage)` = 5,
	Year = 6
  ) %>%
  mutate(
	Number = ifelse(Number %in% c("N/A"), NA, Number),
	Rate = ifelse(Rate %in% c("N/A"), NA, Rate),
	Ratio = ifelse(Ratio %in% c("N/A"), NA, Ratio),
	`Out-of-State (number and percentage)` = 
	  ifelse(`Out-of-State (number and percentage)` %in% 
	           c("N/A"), NA, `Out-of-State (number and percentage)`)
  )  %>%
  filter(!(Number %in% c("25,619",
                     	"30,555",
                     	"35,398",
                     	"37,523",
                     	"73,815",
                     	"89,469",
                     	"59,854",
                     	"27,471"))) %>%
  mutate(Number = as.numeric(gsub(",", "", Number)),
     	Rate = as.numeric(gsub(",", "", Rate)),
     	Ratio = as.numeric(gsub(",", "", Ratio)),
     	Year = as.numeric(gsub(",", "", Year))
     	) %>%
  mutate(Area = tolower(Area)) %>%
  mutate(Abortions_per_1000_women = Rate) %>%
  mutate(Abortions_per_1000_live_births = Ratio) %>%
  mutate(`Out_of_state_abortions(number and percentage)` = `Out-of-State (number and percentage)`) %>%
  select(Area, Abortions_per_1000_women, Abortions_per_1000_live_births, Year, `Out_of_state_abortions(number and percentage)`)

head(state_abortion_data)
```

```{r}
state_immigration_data <- state_immigration_data %>%
  slice(-(1:6)) %>%
  rename(
    `Current residence in`  = 1,
    `Population 1 year and over: Estimate` = 2,
    `Population 1 year and over: MOE` = 3,
    `Same house 1 year ago: Estimate` = 4,
    `Same house 1 year ago: MOE` = 5,
    `Same state of residence 1 year ago: Estimate` = 6,
    `Same state of residence 1 year ago: MOE` = 7,
    `Different state of residence 1 year ago: Estimate` = 8,
    `Different state of residence 1 year ago: MOE` = 9,
    Year = 10
  ) %>%
  slice(-(1:6)) %>%
  mutate(
    `Population 1 year and over: Estimate` = as.numeric(
      gsub(",", "", `Population 1 year and over: Estimate`)
      ),
    `Same house 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Same house 1 year ago: Estimate`)
      ), 
    `Same state of residence 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Same state of residence 1 year ago: Estimate`)
      ), 
    `Different state of residence 1 year ago: Estimate` = as.numeric(
      gsub(",", "", `Different state of residence 1 year ago: Estimate`)
      )
    ) 

head(state_immigration_data)
```

```{r}
state_economic_data <- state_economic_data %>%
  rename(
    State  = 1,
    `Median Household Income` = 2,
    Year = 3
  ) %>%
  mutate(
    State = ifelse(State == "", NA, State),
    `Median Household Income` = 
      ifelse(`Median Household Income` == "", NA, `Median Household Income`)
    ) %>%
  filter(complete.cases(.)) %>%
  mutate(`Median Household Income` = 
           as.numeric(gsub(",", "", `Median Household Income`))) %>%
  mutate(State = tolower(State))

head(state_economic_data)
```

```{r}
presidential_election_data <- presidential_election_data %>%
  subset(select = -c(X1, X, TOTAL.VOTES, TOTAL.VOTES..)) %>%
  mutate(
    FEC.ID = ifelse(FEC.ID == "n/a", NA, FEC.ID),
    `GENERAL.RESULTS` = as.numeric(gsub(",", "", `GENERAL.RESULTS`)),
    STATE = tolower(STATE),
    `R Share of General` = as.numeric(gsub("%", "", `GENERAL..`)) / 100
  ) %>%
  filter(PARTY == "R") %>%
  select(STATE, `R Share of General`, YEAR)

head(presidential_election_data)
```

```{r}
abortion_and_economic <- full_join(state_abortion_data, 
             state_economic_data,
             by = c("Area" = "State",
                    "Year" = "Year"))

abortion_and_economic <- abortion_and_economic %>%
  mutate(Year = as.numeric(gsub(",", "", Year)))

state_immigration_data <- state_immigration_data %>%
  mutate(Year = as.numeric(gsub(",", "", Year))) 

culture_president <- left_join(presidential_election_data, 
                               political_culture_elazar,
                               by = c("STATE"= "state")) 

everything_but_immigration <- left_join(culture_president, 
                                        abortion_and_economic,
                                        by = c("STATE" = "Area",
                                               "YEAR" = "Year"))

all_variables_dataset<- full_join(everything_but_immigration, 
                                  state_immigration_data, 
                                  by = c("STATE" = "Current residence in", 
                                         "YEAR" = "Year"))

head(all_variables_dataset)
```

## Reason for Choosing this Dataset

*High level goal:* To gather data from online repositories regarding immigration and immigration policy, and use it to create a spacial-temporal visualization illustrating our topic of interest in a clear and concise fashion.

We are interested in analyzing how interstate migration is affected by different social, cultural, and political factors. We chose to focus on looking at each state's abortion policy, economic status, political culture, and presidential vote in select years to see if patterns in interstate migration emerge.

*Descriptions of Datasets*: There are several datasets used in the project.

The dataset "state_immigration_data," from the US Census, describes the population estimates and margin of errors of populations, per state and year, of people living in the state one year ago, living in the state in the same house one year ago, living in a different state one year ago, and total population. Recorded are the years 2012, 2016, and 2020. This dataset lets us track the population levels of each state over time, based on who lived where one year before being recorded.

The dataset "state_abortion_data," from the CDC, contains information on the number of abortions per 1000 women in each state in the years 2008, 2012, 2016, and 2020. It also includes information on abortions per 1000 live births and the number of out-of-state abortions. This lets us visualize an important social factor that might motivate people to migrate in or out of their state of residence.

The dataset "presidential_election_data," from the Federal Election Commission, contains the share of Republican votes in each state per election in 2008, 2012, and 2020. This lets us visualize the political demographics in states, which might motivate interstate migration.

The dataset "state_economic_data," from the Federal Reserve, contains information on median household income for each state in the years 2008, 2012, 2016, and 2020. This lets us see if there might be economic reasons for interstate migration, in correlation with the other datasets.

The dataset "political_culture_elazar" correlates each state with a culture tag, as theorized by Daniel Elazar in 1966. This will allow us to recognize patterns between state culture and the other variables of our analysis.

The other datasets combine the above variables of interest in different ways for the purpose of building a predictive model.

## Questions and Analysis Plan

How is interstate migration affected by state abortion policy, economic status, political culture, and presidential vote?

## Weekly Plan of Attack

*Week 10* - Submit proposal.qmd, research changes in sanctuary state policy between 2013-2022 in the states we've chosen, for the purpose of visualizing.

*Week 11* - Wrangle and tidy dataset, will include removing rows, using a handful of pivot_longer functions, etc. Update proposal and project to include feedback from peers. We will work to create visualizations tied to geospatial data. The sanctuary state visualization will also be conducted. Outline the writeup.

*Week 12* - Complete work on the writeup, which will go in the index.qmd file. This will be done by all group members. Begin work on presentation.qmd.

*Week 13* - Touch up the presentation.qmd file, practice the presentation to ensure it fits within time constraints.
